A Match Made in Landfills?

### 2023-03-20 5 Manuscript/manuscript/Manuscript.qmd
RF and SL troubleshooting session
Questions for Office Hours

*Study Area* — This study took place in Prado Herrero, a private cattle ranch located northwest of metropolitan Madrid and is surrounded by agriculture (e.g. beef cattle, cereal grains, legumes, and forage plants (“Renta Agraria,” 2019)). Prado Herrero is located within a nationally protected area (Cuenca Alta del Manzanares Regional Park) and is just north of Santillana Reservoir. This reservoir that was declared an Important Bird Area by Regional Catalogue of Reservoirs and Wetlands of the Community of Madrid due to the numerous resident and migratory species that utilize this water source <CITE>. This cattle ranch has supported a productive white stork rookery where storks have been banded and monitored by biologists at UCM for over 20 years <CITE: Aguirre and Atienza>. During the 2020 to 2021 breeding season, between the months of March to June, stork nests were identified, marked, and monitored for productivity. Of marked nests between the 2020 and 2021 breeding seasons, 31 with banded adults were used both years to lay eggs successfully. All 31 nests were located in ash trees (Fraxinus angustifolius ) found within the cattle pasture. Storks that breed within the Prado Herrero rookery are known to utilize Colmenar Viejo Landfill which is located approximately 12km southeast. Colmenar Viejo is an open-air landfill and it is second largest of it’s kind in the Madrid region <CITE: Alejandro 2022>.

*Sample Collection* — Between March to May 2020 and 2021, we collected feces from marked nests with banded adults in known breeding pairs at 3 points of the breeding season; (1) an adult sample during incubation, (2) an early juvenile sample during the first two weeks of the chicks life when adults are believed to forage on natural sources  <CITE: >, and (3) a late juvenile sample after chicks were past two weeks of age when adults forage on anthropogenic resources. Nests were visited in the late mornings and approximately one gram of fresh feces was collected from the perimeter of the nest structure into a sterile Eppendorf tube (). Samples were maintained cold in a portable cooler with frozen gel packs and frozen in a −20°C freezer within 4 hours of collection and processed at a later date.

@@

### Molecular analysis of ARGs
We performed total DNA extraction directly from fecal samples, by using a pressure filtration technique (QuickGene DNA Tissue Kit S, Fujifilm, Japan) following the manufacturer’s instructions. The 16S rRNA gene was amplified in each DNA sample by real time PCR (rtPCR) in 10-fold dilutions of extracted samples, according to Jiang et al. (2013)<CITE:Fix>. A DNA sample was considered validated when a ten-fold dilution showed a cycle threshold (Ct) less than 25 (Esperón et al., 2018). Once validated, we analyzed DNA samples by with a panel of 21 different ARGs encoding resistance to eight different antimicrobial classes: tetracyclines (tet(A), tet(B), tet(Y), tet(K), tet(M), tet(Q), tet(S), and tet(W)), sulfonamides (sulI and sulII), aminoglycosides (str and aadA), phenicols (catI and catII), macrolides (ermB and ermF), quinolones (qnrS and qnrB), betalactams (blaTEM and mecA), and polymyxins (mcr-1). All rtPCR reactions utilized premade gelled format 96-well plates (Biotools, B &M Labs, S.A., Madrid, Spain), with the exception of blaTEM and mecA genes which used the Sybr GreenTM and TaqManTM probe, respectively. Our thermal cycle was the same for all the rtPCR reactions [6′ 95 °C, 40× (10″ 95 °C, 30″ 60 °C)], with alignment and extension in the same step, at constant temperature of 60 °C. A melting curve step was performed at the end of the qPCR reaction to validate the authenticity of the positive (Nieto-Claudin et al., 2019). We quantified the relative burden of each gene for each sample via the cycle threshold (Ct) for the 16S rRNA and the Ct value of the individual ARG using a previously published formula (Esperón et al., 2020).

## Data import and cleaning

@@

## Exploratory/Descriptive analysis

@@

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

@@

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
ModelAIC1 = readRDS(here("4 Results", "Tables", "ModelAIC1.rds"))
knitr::kable(ModelAIC1)

>>>>

*Study Area* --- This study took place in Prado Herrero, a private cattle ranch located northwest of metropolitan Madrid and is surrounded by agriculture (e.g. beef cattle, cereal grains, legumes, and forage plants ("Renta Agraria," 2019)). Prado Herrero is located within a nationally protected area (Cuenca Alta del Manzanares Regional Park) and is just north of Santillana Reservoir. This reservoir that was declared an Important Bird Area by Regional Catalogue of Reservoirs and Wetlands of the Community of Madrid due to the numerous resident and migratory species that utilize this water source <CITE>. This cattle ranch has supported a productive white stork rookery where storks have been banded and monitored by biologists at UCM for over 20 years \<CITE: Aguirre and Atienza\>. During the 2020 to 2021 breeding season, between the months of March to June, stork nests were identified, marked, and monitored for productivity. Of marked nests between the 2020 and 2021 breeding seasons, 31 with banded adults were used both years to lay eggs successfully. All 31 nests were located in ash trees (Fraxinus angustifolius ) found within the cattle pasture. Storks that breed within the Prado Herrero rookery are known to utilize Colmenar Viejo Landfill which is located approximately 12km southeast. Colmenar Viejo is an open-air landfill and it is second largest of it's kind in the Madrid region \<CITE: Alejandro 2022\>.

*Sample Collection* --- Between March to May 2020 and 2021, we collected feces from marked nests with banded adults in known breeding pairs at 3 points of the breeding season; (1) an adult sample during incubation, (2) an early juvenile sample during the first two weeks of the chicks life when adults are believed to forage on natural sources \<CITE: \>, and (3) a late juvenile sample after chicks were past two weeks of age when adults forage on anthropogenic resources. Nests were visited in the late mornings and approximately one gram of fresh feces was collected from the perimeter of the nest structure into a sterile Eppendorf tube (). Samples were maintained cold in a portable cooler with frozen gel packs and frozen in a −20°C freezer within 4 hours of collection and processed at a later date.

@@

### Molecular analysis of ARGs

We performed total DNA extraction directly from fecal samples, by using a pressure filtration technique (QuickGene DNA Tissue Kit S, Fujifilm, Japan) following the manufacturer's instructions. The 16S rRNA gene was amplified in each DNA sample by real time PCR (rtPCR) in 10-fold dilutions of extracted samples, according to Jiang et al. (2013)<CITE:Fix>. A DNA sample was considered validated when a ten-fold dilution showed a cycle threshold (Ct) less than 25 (Esperón et al., 2018). Once validated, we analyzed DNA samples by with a panel of 21 different ARGs encoding resistance to eight different antimicrobial classes: tetracyclines (tet(A), tet(B), tet(Y), tet(K), tet(M), tet(Q), tet(S), and tet(W)), sulfonamides (sulI and sulII), aminoglycosides (str and aadA), phenicols (catI and catII), macrolides (ermB and ermF), quinolones (qnrS and qnrB), betalactams (blaTEM and mecA), and polymyxins (mcr-1). All rtPCR reactions utilized premade gelled format 96-well plates (Biotools, B &M Labs, S.A., Madrid, Spain), with the exception of blaTEM and mecA genes which used the Sybr GreenTM and TaqManTM probe, respectively. Our thermal cycle was the same for all the rtPCR reactions \[6′ 95 °C, 40× (10″ 95 °C, 30″ 60 °C)\], with alignment and extension in the same step, at constant temperature of 60 °C. A melting curve step was performed at the end of the qPCR reaction to validate the authenticity of the positive (Nieto-Claudin et al., 2019). We quantified the relative burden of each gene for each sample via the cycle threshold (Ct) for the 16S rRNA and the Ct value of the individual ARG using a previously published formula (Esperón et al., 2020).

## Data import and cleaning (Reference file and give details HERE. Reference a ReadME/doc here)

@@

## Exploratory/Descriptive analysis (SUMMARY STATS IN EXPLORATION FILE TBD -\> Burden/MDR associated with each class. summary table? with actual summary stats)

@@

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
#resulttable=readRDS("../../results/summarytable.rds")
#knitr::kable(resulttable)
```

@@

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
ModelAIC1 = readRDS(here("4 Results", "Tables", "ModelAIC1.rds"))
ModelAIC1
knitr::kable(ModelAIC1)
